Understanding Structure and Selectivity in Nanostructured Lyotropic Liquid Crystal Membranes using Molecular Simulations

### 2020-04-13 Ben_Manuscripts/hdphmm/Supporting_Information.tex
Added missing intro figs

  \begin{enumerate}
  	\item Parameterize in $x$, $y$, $z$ coordinates with $x$, $y$ coordinates relative to the nearest
  	pore center (see Figure~\ref{fig:xyz_hmm}). 	
  	\item Convert x and y coordinates to radial coordinates, $r$, then fix the state sequence
  	and use the inference component of the IHMM in order to estimate the parameters in terms
  	of $r$ and $z$ (see Figure~\ref{fig:rz_hmm}). 	 	
  	\item Cluster the VAR parameters from the states found in all 24 trajectories. Reassign the state
  	sequence so that segments which belong to the same cluster are labeled the same across all solute
  	trajectories.
  	\item Subtract the mean of each segment, obtained before clustering, in order to zero out the
  	trajectory.
  	\item Fix clustered state sequence and infer parameters of each state, assuming a mean of zero
  	for all states (see Figure~\ref{fig:zeroed_hmm}). 
  \end{enumerate}

  \begin{figure}
    \centering
	\begin{subfigure}{0.8\textwidth}
		\includegraphics[width=\textwidth]{xyz_hmm.pdf}
		\caption{}\label{fig:xyz_hmm}
	\end{subfigure}  	
	\begin{subfigure}{0.8\textwidth}
		\includegraphics[width=\textwidth]{rz_hmm.pdf}
		\caption{}\label{fig:rz_hmm}
	\end{subfigure}  
	\begin{subfigure}{0.8\textwidth}
		\includegraphics[width=\textwidth]{zeroed_hmm.pdf}
		\caption{}\label{fig:zeroed_hmm}
	\end{subfigure}  
	\caption{IHMM parameterization is a multi-step procedure. In the above plots, distinct colors
	correspond to distinct state behavior. (a) First we parameterize the trajectories in terms of
	the $x$, $y$, $z$ solute center-of-mass coordinates with the $x$ and $y$ coordinates relative
	to the nearest pore center (b) Next, we convert the $x$ and $y$ 	
	coordinates to radial coordinates, $r$, then fix the state sequence and use the inference 
	component of the IHMM in order to estimate the parameters in terms of $r$ and $z$. Note that
	the segment colors in (b) are the same as (a) because we fixed the state sequence.
	(c) After clustering the VAR parameters and reassigning the state sequence so that segments 
	which belong to the same cluster are labeled the same across all solute trajectories, we subtract
	the mean of each segment, obtained before clustering, in order to zero out the
  	trajectory, and then fix the clustered state sequence to infer the parameters of each clustered 
  	state, assuming a mean of zero for all states. Note that the colors are no longer aligned with 
  	figures (a) and (b) because the states are relabeled in terms of their clusters. There are
  	also significantly less states in this single trajectory.
	}\label{fig:hmm_demo}
  \end{figure}

  Given enough data, parameter estimates converge within only a few steps. In 
  Figure~\ref{fig:convergence}, we plot the entries of $A$ and $\Sigma$ as 
  a function of IHMM iteration. Some states are sampled far more frequently
  than others. High sampling frequently results in higher certainty converged
  parameter estimates.

  \begin{figure}[h]
  \centering
  \begin{subfigure}{0.7\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_4.pdf}
  \caption{15 total emissions}\label{fig:convergence_MET_low}
  \end{subfigure}
  \begin{subfigure}{0.7\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_21.pdf}
  \caption{1289 total emissions}\label{fig:convergence_MET_medium}
  \end{subfigure}
  \begin{subfigure}{0.7\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_15.pdf}
  \caption{24468 total emissions}\label{fig:convergence_MET_high}
  \end{subfigure}
  \caption{As the total number of emissions observed from each state
  increases, the uncertainty in the converged parameters estimates
  decreases.}\label{fig:fixed_state_convergence}
  \end{figure}

@@

  The states identified by the IHMM are heavily influenced by the Gaussian prior
  placed on $\mathbf{c}$ in Equation~\ref{M-eqn:var} of the main text. The 
  entries of $A$ and $\Sigma$ do not vary over a wide range, so the final
  parameters were relatively insensitive to the priors. In order to 
  maximally automate the IHMM procedure, we attempted to parameterize the
  prior on $\mathbf{c}$ in an intelligent way. The prior parameters should
  be chosen such that the mean level of each state lies within a region of 
  reasonable probability of the prior (see Figure~\ref{fig:prior_guesses}).
  In each dimension, we defined the prior mean to lie halfway between the 
  maximum and minimum of each trajectory dimension. To parameterize the 
  prior's variance in each dimension, we defined the maximum and minimum to 
  be 2 standard deviations from the prior mean. Although this approach has 
  worked quite well for the data in this work, it is important to check the 
  results to determine whether further adjustments to the prior might be needed.
  In Figure TBD, we show the result of a parameterization where the prior
  parameters of $c$ were poorly chosen. %TODO

  \begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{prior_guesses.pdf}
  \caption{The parameters of the prior on the mean vector, $c$, (black line), which
  represents the coordinates at which solute are trapped, should be chosen such
  that the mean levels of each state identified in the trajectory (blue line) lie within
  regions of the prior with reasonable probability. We chose the mean of the prior 
  as halfway between the maximum and minimum of each trajectory dimension. We chose 
  $\sigma$ of the prior by defining the maximum and minimum to be 2 standard deviations
  from the mean (dashed lines).}\label{fig:prior_guesses}
  \end{figure}

@@

  \begin{figure}[h]
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{silhouette_MET.pdf}
  \caption{methanol}\label{fig:silhouette_MET}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{silhouette_URE.pdf}
  \caption{urea}\label{fig:silhouette_URE}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{silhouette_GCL.pdf}
  \caption{ethylene glycol}\label{fig:silhouette_GCL}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{silhouette_ACH.pdf}
  \caption{acetic acid}\label{fig:silhouette_ACH}
  \end{subfigure}
  \caption{We used the silhouette score test in order to quantitatively
  inform our decision on the numbrer of clusters. Silhouette scores 
  closer to one indicate good clustering. The crosses on the plot indicate
  our choice in number of clusters after careful consideration of the
  silhouette scores in conjunction with qualitative examination of the clusters
  formed by the choice. Although the test often
  favors fewer states, we elected use intermediate number of states in 
  the cases of $A$ and $\Sigma$ because choosing 2--3 clusters 
  insufficiently separates the dynamical modes. }\label{fig:silhouette}
  \end{figure}

>>>>

  \begin{enumerate}
  	\item Parameterize in $x$, $y$, $z$ coordinates with $x$, $y$ coordinates relative to the nearest
  	pore center (see Figure~\ref{fig:xyz_hmm}). 	
  	\item Convert $x$ and y coordinates to radial coordinates, $r$, then fix the state sequence
  	and use the inference component of the IHMM in order to estimate the parameters in terms
  	of $r$ and $z$ (see Figure~\ref{fig:rz_hmm}). 	 	
  	\item Cluster the VAR parameters from the states found in all 24 trajectories. Reassign the state
  	sequence so that segments which belong to the same cluster are labeled the same across all solute
  	trajectories.
  	\item Subtract the mean of each segment, obtained before clustering, in order to zero out the
  	trajectory.
  	\item Fix clustered state sequence and infer parameters of each state, assuming a mean of zero
  	for all states (see Figure~\ref{fig:zeroed_hmm}). 
  \end{enumerate}

  \begin{figure}
    \centering
	\begin{subfigure}{0.8\textwidth}
		\includegraphics[width=\textwidth]{xyz_hmm.pdf}
		\caption{}\label{fig:xyz_hmm}
	\end{subfigure}  	
	\begin{subfigure}{0.8\textwidth}
		\includegraphics[width=\textwidth]{rz_hmm.pdf}
		\caption{}\label{fig:rz_hmm}
	\end{subfigure}  
	\begin{subfigure}{0.8\textwidth}
		\includegraphics[width=\textwidth]{zeroed_hmm.pdf}
		\caption{}\label{fig:zeroed_hmm}
	\end{subfigure}  
	\caption{IHMM parameterization is a multi-step procedure. In the above plots, distinct colors
	correspond to distinct state behavior. (a) First we parameterize the trajectories in terms of
	the $x$, $y$, $z$ solute center-of-mass coordinates with the $x$ and $y$ coordinates relative
	to the nearest pore center (b) Next, we convert the $x$ and $y$ 	
	coordinates to radial coordinates, $r$, then fix the state sequence and use the inference 
	component of the IHMM in order to estimate the parameters in terms of $r$ and $z$. Note that
	the segment colors in (b) are the same as (a) because we fixed the state sequence.
	(c) After clustering the VAR parameters and reassigning the state sequence so that segments 
	which belong to the same cluster are labeled the same across all solute trajectories, we subtract
	the mean of each segment, obtained before clustering, in order to zero out the
  	trajectory, and then fix the clustered state sequence to infer the parameters of each clustered 
  	state, assuming a mean of zero for all states. Note that the colors are no longer aligned with 
  	figures (a) and (b) because the states are relabeled in terms of their clusters. There are
  	also less states in this single trajectory.
	}\label{fig:hmm_demo}
  \end{figure}

@@

  The states identified by the IHMM are heavily influenced by the Gaussian prior
  placed on $\mathbf{c}$ in Equation~\ref{M-eqn:var} of the main text. The 
  entries of $A$ and $\Sigma$ do not vary over a wide range, so the final
  parameters were relatively insensitive to the priors. In order to 
  maximally automate the IHMM procedure, we attempted to parameterize the
  prior on $\mathbf{c}$ in an intelligent way. The prior parameters should
  be chosen such that the mean level of each state lies within a region of 
  reasonable probability of the prior (see Figure~\ref{fig:prior_guesses}).
  In each dimension, we defined the prior mean to lie halfway between the 
  maximum and minimum of each trajectory dimension. To parameterize the 
  prior's variance in each dimension, we defined the maximum and minimum to 
  be 2 standard deviations from the prior mean. Although this approach has 
  worked quite well for the data in this work, it is important to check the 
  results to determine whether further adjustments to the prior might be needed.
%  In Figure TBD, we show the result of a parameterization where the prior
%  parameters of $c$ were poorly chosen. %TODO

  \begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{prior_guesses.pdf}
  \caption{The parameters of the prior on the mean vector, $\mathbf{c}$, (black line), which
  represents the coordinates at which solute are trapped, should be chosen such
  that the mean levels of each state identified in the trajectory (blue line) lie within
  regions of the prior with reasonable probability. We chose the mean of the prior 
  as halfway between the maximum and minimum (shown by the dashed lines) of each trajectory dimension. We chose 
  $\sigma$ of the prior by defining the maximum and minimum to be 2 standard deviations
  from the mean.}\label{fig:prior_guesses}
  \end{figure}

  \pagebreak

  The state sequence appears to converge within 500--1000 IHMM iterations. Step 1
  in the procedure from the previous section is the only time the state sequence 
  is determined. We ran 2000 iterations of the IHMM algorithm in order to arrive
  at a finalized state sequence. Stabilization of the VAR parameters indicates 
  that the state sequence has converged. In Figure~\ref{fig:convergence3d}, we plot
  the diagonal entries of 3 different $A$ and $\Sigma$ matrices as well as the 
  entries of the mean vector $\mathbf{c}$, fit to a single 3D methanol trajectory.
  States with a relatively small number of observations are more highly influenced 
  by the boundaries of the state segments and generally lead to much higher 
  variance of the $A$ and $\Sigma$ parameters. In these cases, the mean is
  a more reliable indicator of convergence since stabilized means imply that the
  fluctuations assigned to each state come from motion about a static location.

  \begin{figure}[h]
  \centering
  \begin{subfigure}{0.8\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_47.pdf}
  \caption{706 total emissions}\label{fig:convergence3d_MET_low}
  \end{subfigure}
  \begin{subfigure}{0.8\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_36.pdf}
  \caption{1000 total emissions}\label{fig:convergence3d_MET_medium}
  \end{subfigure}
  \begin{subfigure}{0.8\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_59.pdf}
  \caption{201 total emissions}\label{fig:convergence3d_MET_high}
  \end{subfigure}
  \caption{Given a sufficient number of observations for a given mode, the VAR parameters
  appear to converge within 500 -- 1000 iterations. States with too few observations
  are more highly influenced by the boundaries of the state segments and generally
  lead to much higher fluctuations in the VAR parameters.}\label{fig:convergence3d}
  \end{figure}

  \clearpage

  When we fix the state sequence in subsequent steps of the procedure, parameter 
  estimates converge within only a few steps. In Figure~\ref{fig:fixed_state_convergence}, we 
  plot the entries of $A$ and $\Sigma$ as a function of IHMM iteration. Some states
  are sampled far more frequently than others. High sampling frequently results in
  higher certainty converged parameter estimates. Due to fast convergence, we only
  run 100 iterations of the inference behavior when the state sequence is fixed.

  \begin{figure}[h]
  \centering
  \begin{subfigure}{0.6\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_4.pdf}
  \caption{15 total emissions}\label{fig:convergence_MET_low}
  \end{subfigure}
  \begin{subfigure}{0.6\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_21.pdf}
  \caption{1289 total emissions}\label{fig:convergence_MET_medium}
  \end{subfigure}
  \begin{subfigure}{0.6\textwidth}
  \includegraphics[width=\textwidth]{convergence_MET_15.pdf}
  \caption{24468 total emissions}\label{fig:convergence_MET_high}
  \end{subfigure}
  \caption{The VAR parameters converge quickly when we fix the state
  sequence. As the total number of emissions observed from each state
  increases, the uncertainty in the converged parameters estimates
  decreases.}\label{fig:fixed_state_convergence}
  \end{figure}

  \clearpage

@@

  \begin{figure}[h]
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{silhouette_MET.pdf}
  \caption{methanol}\label{fig:silhouette_MET}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{silhouette_URE.pdf}
  \caption{urea}\label{fig:silhouette_URE}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{silhouette_GCL.pdf}
  \caption{ethylene glycol}\label{fig:silhouette_GCL}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{silhouette_ACH.pdf}
  \caption{acetic acid}\label{fig:silhouette_ACH}
  \end{subfigure}
  \caption{We used the silhouette score test in order to quantitatively
  inform our decision on the numbrer of clusters. Silhouette scores 
  closer to one indicate good clustering. The crosses on the plot indicate
  our choice in number of clusters after careful consideration of the
  silhouette scores in conjunction with qualitative examination of the clusters
  formed by the choice. Although the test often
  favors fewer states, we elected to use an intermediate number of states in 
  the cases of $A$ and $\Sigma$ because choosing 2--3 clusters 
  insufficiently separates the dynamical modes. }\label{fig:silhouette}
  \end{figure}

### 2020-04-13 Ben_Manuscripts/thesis/structure.tex
Added missing intro figs

  \begin{figure}[!htb]
  \centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=0.6\textwidth]{cartoon_pore.pdf}
		\caption{}~\label{fig:undetailed_pore}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=0.6\textwidth]{detailed_pore.pdf}
		\caption{}~\label{fig:detailed_pore}
	\end{subfigure} 
    \caption{(a) Previous understanding of the pores are essentially speculations 
    based on limited chemical and experimental data. (b) We use detailed molecular 
    modeling in this paper in order to appropriately model the pore's complex architecture,
    which is crucial to understanding the mechanism of solute transport. In both 
    pictures, the head group region is colored blue and the tail region is colored cyan.}~\label{fig:detail}
  \end{figure}

  In this chapter, we perform molecular modeling of the Col\textsubscript{h}
  thermotropic (i.e., solvent-free) assembly formed by Na-GA3C11. Compared to the
  lyotropic (i.e, solvent-containing) H\textsubscript{II} phase, the
  Col\textsubscript{h} phase is a simpler starting point. The system is not
  assembled in aqueous phase which allowed us to simulate longer timescales by
  omitting the solvent in the simulations, and there exists detailed experimental
  characterization of the fully aligned state, including 2D-WAXS patterns
  (Figure~\ref{fig:WAXS}) that are useful for reconstructing structural data.

>>>>

%  \begin{figure}[!htb]
%  \centering
%	\begin{subfigure}{0.45\linewidth}
%		\centering
%		\includegraphics[width=0.6\textwidth]{cartoon_pore.pdf}
%		\caption{}~\label{fig:undetailed_pore}
%	\end{subfigure}
%	\begin{subfigure}{0.45\linewidth}
%		\centering
%		\includegraphics[width=0.6\textwidth]{detailed_pore.pdf}
%		\caption{}~\label{fig:detailed_pore}
%	\end{subfigure} 
%    \caption{(a) Previous understanding of the pores are essentially speculations 
%    based on limited chemical and experimental data. (b) We use detailed molecular 
%    modeling in this paper in order to appropriately model the pore's complex architecture,
%    which is crucial to understanding the mechanism of solute transport. In both 
%    pictures, the head group region is colored blue and the tail region is colored cyan.}~\label{fig:detail}
%  \end{figure}
% 
  In this chapter, we perform molecular modeling of the Col\textsubscript{h}
  thermotropic (i.e., solvent-free) assembly formed by Na-GA3C11. Compared to the
  lyotropic (i.e, solvent-containing) H\textsubscript{II} phase, the
  Col\textsubscript{h} phase is a simpler starting point. The system is not
  assembled in aqueous phase which allowed us to simulate longer timescales by
  omitting the solvent in the simulations, and there exists detailed experimental
  characterization of the fully aligned state, including 2D-WAXS patterns
  (Figure~\ref{fig:WAXS}) that are useful for reconstructing structural data.

### 2020-04-13 Ben_Manuscripts/thesis/thesis.tex
add comments to abstract.
Added missing intro figs

\usepackage{amssymb}		% to get all AMS symbols
\usepackage{graphicx}		% to insert figures
%\usepackage{hyperref}		% PDF hyperreferences??
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{textcomp} % gets rid of \perthousand and \micro warnings that come from gensymb
\usepackage{gensymb}
\usepackage{subcaption}
\usepackage{tcolorbox}
\usepackage[export]{adjustbox}
\usepackage{listings}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{diagbox}
\usepackage{array}

@@

  Periodic, nanostructured porous polymer membranes made from the cross-linked
  inverted hexagonal phase of self-assembled lyotropic liquid crystals (LLCs) are
  a promising class of materials for selective separations. Our current understanding
  of these systems is not sufficient to design them for specific applications. In 
  this thesis, we use detailed atomistic molecular modeling in order to investigate 
  structure and transport of solutes within LLC polymer membrane.

  We begin by enhancing our understanding of the nanoscopic pore structure in order 
  to ensure we appropriately model the chemical environment in the pores. We use radial
  distribution functions and simulated X-ray diffraction patterns to quantify and 
  understand the major structural features of the model.

  We use our carefully constructed molecular model in order to observe transport of
  water, sodium ions and 20 small polar solutes within the pores of a lyotropic 
  liquid crystal (LLC) membrane using atomistic molecular simulations. We 
  observe that solutes perform intermittent hops between lengthy periods of 
  entrapment and describe how transport of a species is dependent not only on 
  molecular size, but on chemical functionality as well.

  Based on our qualitative understanding of solute transport, we focus the remainder
  of our efforts on the development of three different stochastic models which can be
  used to generate realizations of solute trajectories with qualitative and quantitative 
  dynamics similar to our MD solute trajectories. Mathematically modeling complex 
  transport phenomena at the molecular level can be a powerful tool for connecting 
  microscopic chemically-dependent solute motions that do not follow simple diffusive
  behavior with macroscopic membrane performance.

  First, using anomalous diffusion theory, we show how solute dynamics can be
  modeled as a fractional diffusion process subordinate to a continuous time
  random walk. We parameterize the distribution of dwell times, hop lengths between
  dwells and correlation between hops. Next, we generalize Markov state models, treating
  the configurational states of the system as a Markov process where each state has 
  distinct transport properties. Finally we implement a third approach,
  called the infinite hidden Markov model, which is capable of automatically 
  distinguishing an unknown number of hidden dynamical modes in multidimensional
  solute time series. This model offers a flexibile approach to modeling solute
  motion and helps us to uncover mechanisms with minimal human effort.
  }

\dedication[Dedication]{	% NEVER use \OnePageChapter here.
	To all of the fluffy kitties.
	}

@@

  I would like to acknowledge and thank a number of people whose support and
  guidance has not gone unnoticed. First, I thank my advisor, Michael Shirts, 
  who has been instrumental to my growth as a competent and confident researcher,
  pushing me to do high quality science. I thank the entire Shirts group for 
  their feedback and encouragement as well as for making our lab an enjoyable to work.
  I thank Richard Noble and Douglas Gin for their guidance on past and present 
  experimental research on LLC membranes. I also thank Michael McGrath, Gregory 
  Dwulet and Sarah Dischinger, members of the Gin and Noble groups, for providing
  invaluable experimental perspective. I thank Matthew Glaser for his 
  contributions to my understanding of liquid crystal systems. Our first paper 
  would not have been complete without his insight. I thank Joseph Yelk for his 
  effort creating the X-ray diffraction code on which our first paper relied.
  I thank Christopher Calderon for his guidance on applying Bayesian non-parametric
  techniques to solute time series.
%  I would like to acknowledge my financial support: The ACS Petroleum Research Fund
%  grant \#59814-ND7 and the Graduate Assistance in Areas of National Need (GAANN) 
%  fellowship which is funded by the U.S. Department of Education. 
%
%  % BJC: This acknowledgement is how they want to be cited in papers. Could I
%  % shorten it for this?
%  Finally, I would like to acknowledge my high performance computing resources:
%  Molecular simulations were performed using the Extreme Science and
%  Engineering Discovery Environment (XSEDE), which is supported by National
%  Science Foundation grant number ACI-1548562. Specifically, it used the Bridges
%  system, which is supported by NSF award number ACI-1445606, at the Pittsburgh
%  Supercomputing Center (PSC). This work also utilized the RMACC Summit supercomputer,
%  which is supported by the National Science Foundation (awards ACI-1532235 and
%  ACI-1532236), the University of Colorado Boulder, and Colorado State
%  University. The Summit supercomputer is a joint effort of the University of
%  Colorado Boulder and Colorado State University.

@@

% \LoFisShort	% use this only for 1-page Table of Figures
% \emptyLoF	% use this if there is no List of Figures

% \LoTisShort	% use this only for 1-page Table of Tables
% \emptyLoT	% use this if there is no List of Tables

@@

\graphicspath{{../transport/figures/}{../structure_paper/figures/}{../transport/supporting_figures/}{../stochastic_transport/figures/}{../stochastic_transport/supporting_figures/}}

@@

%\input structure.tex
%\input structure_si.tex

%\input transport.tex
%\input transport_si.tex

%\input stochastic_transport.tex

@@

%%%%%%%%%   then the Appendices, if any   %%%%%%%%%
\appendix
%\input structure_si.tex
%\input transport_si.tex
%\input stochastic_transport_si.tex

>>>>

\usepackage{amssymb}		% to get all AMS symbols
\usepackage{graphicx}		% to insert figures
\usepackage{hyperref}		% PDF hyperreferences??
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{textcomp} % gets rid of \perthousand and \micro warnings that come from gensymb
\usepackage{gensymb}
\usepackage{subcaption}
\usepackage{tcolorbox}
\usepackage[export]{adjustbox}
\usepackage{listings}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{diagbox}
\usepackage{array}

@@

  Periodic, nanostructured porous polymer membranes made from the cross-linked
  inverted hexagonal phase of self-assembled lyotropic liquid crystals (LLCs) are
  a promising class of materials for selective separations. There is a wide 
  design space available for liquid crystal monomer design which may allow one to
  tailor solute interactions within the membrane pores. Our currrent
  understanding of these membrane is based primarily on experimental measurements
  and macroscopic modeling which limits our insight and ability to design them 
  for specific applications.

  In this thesis, I develop a detailed atomistic molecular model, maximally consistent
  with experimental structural data, in order to investigate structure and transport
  of solutes within LLC polymer membranes. I characterize the dynamics of water, sodium
  ions and 20 small polar solutes within its pores. I present evidence which suggests that 
  LLC membranes are characterized by a gradual, rather than abrupt, transition in 
  composition from the primarily aqueous pore region to the hydrophobic monomer
  tail region. Within this continuous and heterogeneous environment, solutes perform
  intermittent hops between lengthy periods of entrapment. I show how transport rates
  of these species are dependent not only on molecular size, but on chemical functionality
  as well.

  In order to connect these microscopic chemically-dependent solute motions to
  macroscopic membrane performance, I develop three different stochastic models 
  of varying complexity.  
  First, using anomalous diffusion theory, we show how solute dynamics can be
  modeled as a fractional diffusion process subordinate to a continuous time
  random walk. We parameterize the distribution of dwell times, hop lengths between
  dwells and correlation between hops. Next, we generalize Markov state models, treating
  the configurational states of the system as a Markov process where each state has 
  distinct transport properties. Finally apply the infinite hidden Markov model, 
  an approach capable of automatically distinguishing and parameterizing an unknown 
  number of hidden dynamical modes in multidimensional solute time series. This model
  offers a flexibile approach to modeling solute motion and helps us to uncover 
  mechanisms with minimal human effort.

\dedication[Dedication]{	% NEVER use \OnePageChapter here.
	To my girls Jenna, Maclynn and Aria.
	}

@@

  I would like to acknowledge and thank a number of people whose support and
  guidance has not gone unnoticed. First, I thank my advisor, Michael Shirts, 
  who has been instrumental to my growth as a competent and confident researcher,
  pushing me to do high quality science. I thank the entire Shirts group for 
  their feedback and encouragement as well as for making our lab an enjoyable
  place to work. I thank Richard Noble and Douglas Gin for their guidance on past
  and present experimental research on LLC membranes as well as their
  students Michael McGrath, Gregory Dwulet and Sarah Dischinger, for providing 
  invaluable experimental perspective. I thank
  Matthew Glaser for his contributions to my understanding of liquid crystal systems.
  Our first paper would not have been complete without his insight. I thank Joseph
  Yelk for his effort creating the X-ray diffraction code on which our first paper
  relied. Finally, I thank Christopher Calderon for his guidance on applying 
  Bayesian non-parametric techniques to solute time series, allowing a leap in
  the flexibility of our dynamical modeling approaches.

  Outside of the University, I was fortunate enough to have full support of my
  friends and family. Moving to Colorado was a big leap and I would not
  have been successful without their encouragement.

@@

% \LoFisShort	% use this only for 1-page Table of Figures
\emptyLoF	% use this if there is no List of Figures

% \LoTisShort	% use this only for 1-page Table of Tables
\emptyLoT	% use this if there is no List of Tables

@@

\graphicspath{{../transport/figures/}{../structure_paper/figures/}{../transport/supporting_figures/}{../stochastic_transport/figures/}{../stochastic_transport/supporting_figures/}{../hdphmm/figures/}{../hdphmm/supporting_figures/}}

@@

\input structure.tex
%\input structure_si.tex

\input transport.tex
%\input transport_si.tex

\input stochastic_transport.tex

\input hdphmm.tex

@@

%%%%%%%%%   then the Appendices, if any   %%%%%%%%%
\appendix
\input structure_si.tex
\input transport_si.tex
\input stochastic_transport_si.tex
\input hdphmm_si.tex
